---
title: "Homework 2"
author: "Rosemary Kinuthia"
date: "2/20/2018"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(car)
library(tidyverse)
```
#Question 1: What kind of R object is the Davis dataset?
```{r}
#Load Davis dataset
dataDavis <-car::Davis
```

```{r}
#Look at type of object
class(dataDavis)
```
#Question 2: How many observations are in the Davis dataset?
```{r}
#Summary of Data
summary(dataDavis)
```
#Question 3: For reported weight, how many observations have a missing value?
```{r}
#Complete cases
completeObs <- complete.cases(dataDavis)
table(completeObs)
```
#Question 4: How many observations have no missing values? 
```{r}
#Keep only the complete data
dataComplete <- dataDavis %>%
  filter(complete.cases(.))
```
#Question 5: How many females are in this subset?
```{r}
#Create subset containing only sex
dataDavis_subset <- select(dataDavis, 
                           sex)
dim(dataDavis_subset)
```

```{r}
#Create subset containing only females
Female_subset <- filter(dataDavis_subset, 
                    sex == "F")
dim(Female_subset)
```
#Question 6: What is the average BMI for these individuals?
```{r}
#Compute a new variable (Height in Meters)
dataDavis <-dataDavis %>%
  mutate(HeightMeters= height/100)
```

```{r}
#Compute a new variable (Height in Meters Squared)
dataDavis <- dataDavis%>%
  mutate(HeightMeters_Squared= HeightMeters*HeightMeters)
```

```{r}
#Create a new variable (BMI)
dataDavis <- dataDavis%>%
  mutate(BMI=weight/HeightMeters_Squared)
```

```{r}
#Calculate average BMI (Look at summary statistics)
summary(dataDavis)
#Mean BMI is 24.7
```
#Question 7: How do these individuals fall into the BMI categories (what are the frequencies and relative %’s)?
```{r}
#Recode data
dataDavis <- dataDavis %>%
  mutate(BMIcat = if_else (BMI<18.5, "1. Underweight",
                  if_else (BMI<25, 
                          "2. Normal",
                  if_else (BMI<30,  
                          "3. Overweight",
                          "4. Obese"),
                  "4. missing")))
```

```{r}
#Count the number of individuals that fall into the BMI categories 
dataDavis %>%
  count(BMIcat)
```

```{r}
#Formatted table of contents
library(janitor)
dataDavis %>%
  janitor::tabyl(BMIcat) 

dataDavis %>%
  janitor::tabyl(BMIcat) %>%
  knitr::kable()
```
#Question 8: Create a histogram of BMI.
```{r}
#Histogram of BMI
dataDavis%>%
  ggplot()+
  geom_histogram(aes(BMI))+
  ggtitle("Histogram of BMI Data in Davis Dataset")
#I noticed that there is an outlier of BMI around 500. I will delete the outlier below.
```

```{r}
#Remove outlier (BMI of 500)
dataDavis_OutlierDeleted <- dataDavis %>%
  filter (BMI<100)
dim(dataDavis_OutlierDeleted)
```

```{r}
#New histogram of BMI with outlier removed
dataDavis_OutlierDeleted%>%
  ggplot()+
  geom_histogram(aes(BMI))+
  ggtitle("Histogram of BMI Data in Davis Dataset")
#After deleting the outlier I noticed that the distribution is skewed to the right. 
```
#Question 9: Create side-by-side boxplots of the BMI distributions by gender
```{r}
#Side-by-side boxplots of the BMI distribution by gender
ggplot(dataDavis_OutlierDeleted, 
       aes(x=sex, y=BMI, fill=sex)) + 
  geom_boxplot()
```
#Question 10: Create a clustered bar chart of the BMI categories by gender
```{r}
#Clustered barchart of the BMI categories by gender
dataDavis_OutlierDeleted %>% ggplot(aes(x=BMIcat, fill=sex)) +
  geom_bar(position = "dodge")
```

##The link to this assignment can be found at [https://github.com/RosemaryKinuthia/N741Homework2.git] (https://github.com/RosemaryKinuthia/N741Homework2.git)

